% --- English Version ---
Writing this thesis and completing this academic journey has been one of the most challenging yet profoundly rewarding experiences of my life. While this document bears only my name, it is the culmination of the support, guidance, and encouragement of many people to whom I owe my deepest gratitude.

\vspace{1em}
First, I would like to thank my thesis supervisor, Prof. Giovanni Malnati, for their guidance, patience, and expertise throughout this project. I am also deeply grateful to the team at the Lutech company, particularly for providing the infrastructure and collaborative environment that made this research possible.

\vspace{1em}
To my family: thank you for your unconditional love and unwavering support. 

\vspace{1em}
To my parents, thank you for always believing in me, for the sacrifices you have made to give me this opportunity, and for being a constant source of encouragement even during the most stressful phases of my studies. I truly could not have reached this milestone without you.

\vspace{1em}
To my university mates: thank you for the shared struggles, the late-night debugging sessions, and the countless coffees that kept us going. Navigating the complexities of exams, projects, and deadlines was made infinitely easier --- and much more fun --- because we did it together. The friendship we built is something I will always carry with me.

\vspace{1em}
To my friends: thank you for keeping me grounded. Your ability to pull me away from the computer screen, listen to my frustrations, and remind me of the world outside of university has been essential to my sanity and well-being.

\vspace{1em}
Finally, a note of gratitude to myself --- for the perseverance, the long nights, and the refusal to give up in those moments when everything seems to fall apart. This achievement belongs to all of us.

\vspace{1em}
Thank you.

\clearpage

% --- Italian Version ---
\begin{itshape} % Italics help visually separate the two languages beautifully
Scrivere questa tesi e completare questo percorso accademico è stata una delle esperienze più impegnative ma anche più profondamente gratificanti della mia vita. Sebbene questo documento porti solo il mio nome, è il culmine del supporto, della guida e dell'incoraggiamento di molte persone a cui devo la mia più profonda gratitudine.

\vspace{1em}
Innanzitutto, vorrei ringraziare il mio relatore, Prof. Giovanni Malnati, per la sua guida, la sua pazienza e la sua competenza durante tutto questo progetto. Sono anche profondamente grato al team dell'azienda Lutech, in particolare per aver fornito l'infrastruttura e l'ambiente collaborativo che hanno reso possibile questa ricerca.

\vspace{1em}
Alla mia famiglia: grazie per il vostro amore incondizionato e il vostro incrollabile supporto. 

\vspace{1em}
Ai miei genitori, grazie per aver sempre creduto in me, per i sacrifici che avete fatto per darmi questa opportunità e per essere stati una costante fonte di incoraggiamento anche durante le fasi più stressanti dei miei studi. Non avrei davvero potuto raggiungere questo traguardo senza di voi.

\vspace{1em}
Ai miei compagni di università: grazie per le fatiche condivise, le sessioni notturne di debugging e gli innumerevoli caffè che ci hanno dato la forza di andare avanti. Affrontare la complessità di esami, progetti e scadenze è stato infinitamente più facile --- e molto più divertente --- perché lo abbiamo fatto insieme. L'amicizia che abbiamo costruito è qualcosa che porterò sempre con me.

\vspace{1em}
Ai miei amici: grazie per avermi tenuto con i piedi per terra. La vostra capacità di allontanarmi dallo schermo del computer, di ascoltare le mie frustrazioni e di ricordarmi il mondo al di fuori dell'università è stata essenziale per il mio equilibrio e il mio benessere.

\vspace{1em}
Infine, una nota di gratitudine a me stesso --- per la perseveranza, le lunghe notti e il rifiuto di arrendermi in quei momenti in cui tutto sembrava crollare. Questo traguardo appartiene a tutti noi.

\vspace{1em}
Grazie.
\end{itshape}